---
title: "VECM Analysis"
author: "Naim Sejtanic"
output:
  html_document: default
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
library(plyr)
library(magrittr)
library(dplyr)
library(urca)
library(doBy)
library(zoo)
library(rvest)
library(rmarkdown)
library(tinytex)
library(forecast)

knitr::opts_chunk$set(echo = TRUE)
```


## VECM Analysis
In this analysis, we will observe the forward rate and the spot exchange rate and analyze whether or not cointigration exists, and create a VECM forecasting model if it does. We use the data from spot.csv.

First, we test for speculative efficiency by considering the equation $s_t = \beta_0+ \beta_1 f_{t-n}+u_t,$ where $s_t$ is the log spot rate, $f_{t-n}$ is the log forward rate lagged n periods, and $u_t$ is a disturbance term. We will use the unit root test to determine the level of integration of $s_t$ and $f_t$.

```{r}
d0 <- read.csv("spot.csv") 
logspot = log(d0$spot)
log1 = log(d0$forward1)
log3 = log(d0$forward3)
log6 = log(d0$forward6)

dftest_spot = ur.df(y = logspot, type = "drift", selectlags = "AIC")
summary(dftest_spot)

```
```{r}
dftest_forward1 = ur.df(y = log1, type = "drift", selectlags = "AIC")
summary(dftest_forward1)
```

```{r}
dftest_forward3 = ur.df(y = log3, type = "drift", selectlags = "AIC")
summary(dftest_forward3)
```

```{r}
dftest_forward6 = ur.df(y = log6, type = "drift", selectlags = "AIC")
summary(dftest_forward6)
```
In all of the above tests, we fail to reject the null hypothesis since all of the test statistics are insignificant at the 5% test level. We need to construct the first difference of the log rates and repeat the unit root tests. 

```{r}
dlogspot = diff(log(d0$spot))
dlog1 = diff(log(d0$forward1))
dlog3 = diff(log(d0$forward3))
dlog6 = diff(log(d0$forward6))

dftest_spot1 = ur.df(y = dlogspot, type = "drift", selectlags = "AIC")
summary(dftest_spot1)

```

```{r}

dftest_forward1d = ur.df(y = dlog1, type = "drift", selectlags = "AIC")
summary(dftest_forward1d)
```


```{r}
dftest_forward3d = ur.df(y = dlog3, type = "drift", selectlags = "AIC")
summary(dftest_forward3d)

```

```{r}
dftest_forward6d = ur.df(y = dlog6, type = "drift", selectlags = "AIC")
summary(dftest_forward6d)

```
In all the above tests, we reject the null hypothesis since all test statistics are significant at the 5% level. Hence, we conclude that the first difference of log spot rate $\Delta s_t$ and the first difference of forward rate $\Delta f_t$ are stationary time series, since the level of integration of $s_t$ and $f_t$ is 1. 


Next, we test for cointegration between $\gamma_t = (s_t, f_{t-4})$ by specifying the VECM $\Delta \gamma_t = \alpha(\beta'\gamma_{t-1}-\beta_0)+v_t$
```{r}
matplot(y = cbind(logspot, lag(log1,4)), type = "l", xlab = "Time")
```
```{r}
result = lm(logspot ~ lag(log1, 4))
summary(result)

```
We get the estimated equation: $s_t = -0.0224 + 0.9117 f_{t-4}+u_t$.
Now, we can specify the following VECM model: $\Delta \gamma_t = \alpha(s_t-0.9117 f_{t-4}+0.0224)+v_t$ 

We can use Johansen test to check the the null hypothesis that the residual is I(1), which means no cointegration. The alternative hypothesis: residual is I(0), which means cointegration exists. 

Johansen test
```{r}
d1 = data.frame(logspot, lag(log1,4))
jotest<- ca.jo(d1, ecdet = "none", spec = "transitory", K=2)
summary(jotest)

```


Since the test statisitic of r=0 is 140.26, we are able to reject the r=0 null hypothesis, this suggests that $f_{t-4}$ the lagged 1 month forward rate is cointegrated with the $s_t$ spot exchange rate. 


We can now estimate the bivariate system of VECM, which is listed below 

$$\begin{pmatrix}
\Delta s_{t} \\
\Delta f_{t-4} 
\end{pmatrix}
=   
\begin{pmatrix}
\alpha_1 \\
\alpha_2 
\end{pmatrix}
 (s_t-0.9117 f_{t-4}+0.0224)  + 
\begin{pmatrix}
v_{1,t} \\
v_{2,t} 
\end{pmatrix}$$
First, we estimate the equation for $\Delta s_t$, which is $\Delta s_t = \alpha_1(s_t-0.9117 f_{t-4}+0.0224)+v_{1t}$


```{r}
cointegrated = logspot- lag(log1, 4)*summary(result)$coefficients[2,'Estimate']-summary(result)$coefficients[1,'Estimate']
cointegrated = cointegrated[-1]
result1 = lm(dlogspot ~ cointegrated)
summary(result1)

```
$\alpha_1 = 0.2316$ 
Finally, we estimate the equation for $\Delta f_{t-4}$, which is $\Delta f_{t-4} = \alpha_2(s_t-0.9117 f_{t-4}+0.0224)+v_{2t}$

```{r}
result2 = lm(diff(lag(log1,4)) ~ cointegrated)
summary(result2)

```
$\alpha_2 = 0.0380$
Therefore, the estiamted bivariate VECM is 
$$\begin{pmatrix}
\Delta s_{t} \\
\Delta f_{t-4} 
\end{pmatrix}
=   
\begin{pmatrix}
0.2316 \\
0.0380
\end{pmatrix}
 (s_t-0.9117 f_{t-4}+0.0224)$$


